\newpage
\section{Background}

Internet technology is the environment in which billions of people and trillions of
devices are interconnected in various ways.
As part of this evolution, Internet becomes the basic carrier
for interconnecting electronic devices – used in industrial
automation, automotive electronics, telecommunications
equipment, building controls, home automation, medical
instrumentation, etc. – mostly in the same way as the Internet came
to the desktops before. More and more devices getting connected to World Wide
Web. Variety of factors have influenced  this evolutions ~\cite{4221180}:
\begin{itemize}
\item The availability of affordable, high-performance, low-power
electronic components for the consumer devices. Improved technology can assist
building  advanced functionality into embedded devices and enabling new ways of
coupling between them.
\item Even low cost embedded devices have some wired or wireless interface to local area
networks of the Ethernet type. TCP/IP family protocols are becoming the standard
vehicle for exchanging information between networked devices.
\item The emergence of platform independent data interchange mechanisms based on
Extensible Markup Language (XML) data formatting gives lots of opportunities for developing high-level data interchange and
communication standards at the device level.
\item The paradigm of Web Services  helps to connect various
independent applications using lightweight communications. Clients that are
connected to the service and the service itself may be written using different
programming languages and be executed on different platforms.
\item Presence of Internet allows existing of small embedded controllers and
large production servers in the same network, with a possibility to change
information.
\end{itemize}


The integration of different classes of devices, which employ different
networking technologies, is still an open research area. One of the possible
solutions is the use of SOA software architecture design pattern.



\subsection{Service oriented architecture}
\begin{quote}
\textit{
Service-oriented computing is a computing paradigm that uses services as basic building blocks for
application development.}
~\cite {lws_milanovic.pdf}
\end{quote} 

The purpose of \gls{SOA} is to allow easy cooperation of a large number of
computers that are connected over a network.
Every computer can run one or more services, each of them implements one
separate action. This may be a simple business task. Clients can make calls
and receive required data or post some event messages.

Services are self-describing and open components. There is a service interface,
that is based on the exchange of messages and documents using standard formats.
Interface internals ( operating system, hardware platform, programming language)
are hidden from client. Client uses only a service specification scheme, also
called contract. Consumers can get required piece of functionality by mapping problem
solution steps to a service calls. This scheme provides quick access and easy integration of software components.


Service architecture have been successfully adopted in
business environments. Different information systems, that were created inside
companies for automation of business processes,  are now turned into services
which may easily interact with each other. For example, Estonian government uses
services to transmit data between information systems of different departments.
There are also some free services available. Some Internet search companies like
Google, Bing, Yandex provide lots of alternatives how to retrieve data without
using regular browser( search , geolocation and maps,  spell check \gls{API}s )

There are available many technologies which can be used to implement
\gls{SOA}~\cite{wikipedia:SOA}:

\begin{itemize}
  \item \label{itm:soa_technologies} \textbf{Web Services}
  \item{\textbf{SOAP}} Simple Object Access Protocol, is a protocol specification for exchanging structured information in the implementation of Web Services in computer networks.
  \item{\textbf{RPC}} Remote procedure call is an inter-process communication that allows
  a computer program to cause a subroutine or procedure to execute in another address space (commonly on another computer on a shared network) without the programmer explicitly coding the details for this remote interaction. 
  \item{\textbf{REST}} Representational state transfer is a style of software architecture for distributed systems such as the World Wide Web. REST has emerged as a predominant web API design model.
  \item{\textbf{DCOM}} Distributed Component Object Model is a proprietary Microsoft technology for communication among software components distributed across networked computers. 
  \item{\textbf{CORBA}}  Common Object Request Broker Architecture enables separate pieces of software written in different
    languages and running on different computers to work with each other like a single application or set of services. Web services
  \item{\textbf{DDS}} Data Distribution Service for Real-Time Systems (DDS) is an Object
  Management Group (OMG) machine-to-machine middleware standard that aims to
  enable scalable, real-time, dependable, high performance and interoperable data exchanges between publishers and subscribers. 
  \item{\textbf{Java RMI}} Java Remote Method Invocation is a Java API that performs the
  object-oriented equivalent of remote procedure calls (RPC), with support for direct transfer of serialized Java objects and distributed garbage collection.
  \item{\textbf{Jini}}  also called Apache River, is a network architecture for the construction of distributed systems in the form of modular co-operating services.
  \item{\textbf{WCF}} The Windows Communication Foundation (or WCF), previously known as
  "Indigo", is a runtime and a set of APIs (application programming interface)
  in the .NET Framework for building connected, service-oriented applications.
  \item{\textbf{Apache Thrift}} is used as a remote procedure call (RPC) framework and
  was developed at Facebook for "scalable cross-language services development".
  \item \ldots
\end{itemize}

This list can be continued. Most of these technologies are inspired by idea
of \gls{RPC}. An \gls{RPC} is initiated by the client, which sends a request
message to a known remote server to execute a specified procedure with specified
parameters. The remote server sends a response to the client, and the application continues its process.
This idea is described in details in \autoref{sec:rpc}.



Web Services are the most popular technology for implementing
service-oriented software nowadays. Next section will focus on this framework
and on the main features that any \gls{SOA} implementation should have.

\input{background/ws_arch.tex}
\input{background/rest.tex}
\input{background/rpc.tex}
\input{background/serialization.tex}
\input{background/soa_embedded.tex}

% TODO DDDDDDDDDDDDDDDDDDDDDDDDDDD
% Service
% -
% Oriented Architecture can be implemented in different ways. General focus is on
% whatever architecture gets the job done. SOAP and REST have their strengths and weaknesses
% and will be highly suitable to some
% applications and positively terrible for others. The decision of
% which to use depends entirely on the circumstances of the application.






% Main constraints - resources of embedded system.
% Target STM32 device.
% Absence of ethernet.
% Existing implementations require more resources than STM32 or similar has.
% TODO another constraints of embedded systems.

% TODO put is somewhere
 % The benefit of SOA is
% to allow simultaneous use and easy mutual data exchange between programs of different vendors without additional programming or making changes to the services.
 

% DDDDDDDDDDDDDDDDDDDDDDDDDDD



%\input{background/embedded_security.tex}




\subsection{Final target system requirements}
% TODO list of final requirements here. This list is needed to be proved by the
% implementation

All previous sections introduced ideas how various devices can communicate to
each other. Each described technology has its own pros and cons. In general, we
need to choose a technology without any drawbacks or a technology, which is the
smallest evil chosen from list of suitable ones.


Web Services have these advantages:
\begin{itemize}
  \item Service description
  \item Service discovery
  \item Portability and platform independence( XML)
  \item Standardized protocols and message structures.
  \item Ability to transfer messages using other transport protocol than HTTP
\end{itemize}

The RESTful approach enables to model our domain objects as
consistent, hierarchical URLs with predictable operations for \gls{CRUD}(GET,
POST, PUT, DELETE). It is also based on HTTP that comes with standard error
codes, message types( see also \nameref{sec:rest_multiple_resource_types}) and
generally does most of the transport hard work, so we benefit from not
needing to maintain any user-developed protocol and using ready and well defined
technologies. One of the main concepts of REST is that RESTful services are
stateless and do not store session information. This reduces
resource consumption and makes client-server applications more loosely coupled
and scalable.

REST and Web services have different ideas. REST is based on resources and their
representation, while Web Services use messages to send data and call remote
methods between server and client. This technology is based on very simple idea
of \gls{RPC}. RPC has some essential benefits over REST and
WS-* technologies: it does not require traditional transport like HTTP, TCP,
Ethernet or Wi-fi. It can be implemented using any radio link, serial line or
any other suitable transport, that is able to deliver response and request
messages. Web Services in theory could also be transport independent, but most
WS-* tools assume the HTTP(and underlying technologies) as de facto standard
\footnote{SOAP protocol transport independence started from SOAP
1.2~\cite{soap_protocol_spec}}. Support of other underlying protocols for
SOAP(the main messaging protocol in WS-*) should be implemented separately and
most standard tools does not have this, they just use standard HTTP. WS-* has a
huge overhead in doing simple things like implementing small light
controlling service in your room. This is not right tool for that. Simple RPC
would be enough.

\autoref{tbl:service_ideas} defines system requirements and features for service prototype in this
work. These are ideas that were kept in mind while developing of embedded
service system was in progress.

\begin{longtabu} to \textwidth {|X|X|}
	
\hline 
\multicolumn{1}{|l|}{\textbf{Feature}} & 
\multicolumn{1}{l|}{\textbf{Description}} \\ 
\hline 
\endfirsthead

\multicolumn{2}{l}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\hline 
\multicolumn{1}{|l|}{\textbf{Feature}} & 
\multicolumn{1}{l|}{\textbf{Description}} \\ 
\hline  
\endhead

\hline \multicolumn{3}{|r|}{{Continued on next page}} \\ \hline
\endfoot

% \hline \hline
\endlastfoot

		Transport independent solution
		&
		Company did not defined final communication yet.
		This may be one of these: UART, Bluetooth, RF, HTTP.
		In this work i implement only first prototype of such system and 
		requirements are about to be changed in future. 
		Therefore i need to implement a portable solution that could be quickly ported
		to another environment.
		
		\tabularnewline
		\hline
			Service description and service contract &
			Gives overview of all service capabilities. \newline 
			Provides an interface definition language. \newline
			Is a specification( sometimes may be only available one) to the server and
			server interface.
	 
	    \tabularnewline
		\hline	
	    RPC based communication &
	    Our prototype needs to control coffee machine and execute some requested
	    actions. If we had an Ethernet, the design decision would be REST(with all
	    its benefits and philosophy).
	    Otherwise RPC is the most suitable solution here. This system is message
	    oriented. Some standard RPC solution should be used or ported to the
	    embedded device.
	    
	    \tabularnewline
		\hline	
		Lightweight and verbose messages &
		Microcontroller has limited resources. We cannot use huge messages for really
		small request data~\footnotemark.
		Some text based protocol need to be used.
		Binary protocols are evil, it is real hard to understand and debug them.
		
	    \tabularnewline
		\hline
		Simple design &
		Simple is better than complex. \newline
		Not like WS-* \\
		\tabularnewline
		\hline
		Modular architecture &
		System should be divided into separate functional modules.
		It should be easy to change the internal implementation of these modules, without a need for global refactoring.		
		\tabularnewline
		\hline	  

	\caption{Ideas about embedded service}
	\label{tbl:service_ideas}
\end{longtabu}

\footnotetext{internet resource \url{http://www.simple-is-better.org/rpc/} contains a comparison of different message formats)}


