\newpage
\section{Implementation}
\label{sec:implementation}
This section will cover the implementation of embedded service server and other
parts of the system. This system is only a small part of whole service
infrastructure, it does not cover discovery, addressing, authentication and
security and other essential parts of every production service.
This prototype only includes an embedded server and a client application to
demonstrate how technologies and methods already implemented in "big computer
systems" can be adapted to "small" embedded and resource-constrained devices.
Implementing a full application stack of service technologies (for example WS-*
or \gls{DPWS}) needs a lot of human and time resources. You can read
some standards( amount of pages was already mentioned above in section
\nameref{sec:adv_and_disadv_of_ws_standards}) and count how many
human*hours it would take to implement this in embedded system environment with limited amount of resources, low-level
application programming using C language and without ready made and
off-the-shelf software tools and libraries for that kind of systems. In my opinion, one master student
is unable to create a complete server solution only by himself within reasonable
time. The scope of this work requires some more resources, a team with several
members maybe, to accomplish such task.

The solution that is possible to implement during university project like this
is the research about related field and available technologies and a simple
system prototype.
I have implemented this using collected features from literature and web
resources.

This design is not based on any specific transport protocol.
Any physical protocol, that is able to deliver characters or bytes, can be used with this application.
There were no any physical and logical protocol requirements set for this prototype solution and the technologies, that company will use in future, are not specified.
Therefore, it is designed as portable as possible and operates only with character data.
It contains an simple example of data link layer protocol, that encapsulates RPC messages. 
Whole communication is transferred over serial line.


First section below covers the general architecture of implemented service.
Next come details about embedded server, which contain the description about
hardware and software platform used, program architecture and data flow.
There is also an implementation of the client side application library (also
called client stub) below.
The last section here introduces one possible client application for
this service architecture.

\input{implementation/system_arch.tex}

\input{implementation/embedded_server.tex}

\input{implementation/java_service_library.tex}
\input{implementation/android_client.tex}