\clearpage\vspace*{\fill}
\section*{Annotation}
Current work introduces conceptual approaches for implementing an extensible
service oriented client-server application on a small microcontroller.
This is a general-purpose transport and hardware independent embedded server
that uses remote procedure calls as primary communication protocol.
This server looks like remote service that could provide defined functions to
the client.

This paper has a research about available service oriented architecture technologies and reviews them in the context of embedded resource constrained hardware systems.

There is also an implementation of a small embedded service and corresponding client.
Service part is implemented on STM32F1 family ARM Cortex-M3 microcontroller and runs FreeRTOS operating system.
It controls a remote object, that is connected over serial line and uses some closed protocol. 
This server device is a bridge between client and controlled remote object. 
It translates all communication between them and can be used to integrate proprietary or legacy system into existing environment. 

Client part is realized as Java RPC stub library  and the Android demo application, which is capable to execute some functions on the remote object.
An example of controlled device here is the coffee machine, which can prepare a cup of coffee for remote client.

Client and server are communicating using lightweight JSON-RPC protocol over Bluetooth wireless communication line.
Some few RPC call functions are implemented to show the capabilities of such client-server architecture.

This work contains several possibilities how to organize communication between different embedded systems.

\vspace{\fill}
\clearpage